A Geometry for Network Visualization in \MakeLowercase{\texttt{ggplot2}}

### 2015-03-30 CC/writeup.Rnw
fix ex.
changed email example ... I like it!

\subsection{Email Network}
  \emph{Note: would be best with coloring and direction}
  This email network comes from the 2014 VAST Challenge \citep{emailnet}. It is a directed network of emails between company employees with 55 vertices and 9,063 edges. Each vertex is an employee of the company, and each edge is an email sent from one employee to one or more other employees. The arrow of the directed edge points to the recipient(s) of the email. The network contains two business weeks of emails across the entire company. }
<<email_ex, fig.cap = 'All emails within a company over a two week period.'>>=
ggplot(data = email.net, aes(from_id = From, to_id = to)) +
  geom_net(vertices = email.names,ealpha = 0.1, vsize = I(2), directed = TRUE) +
  expand_limits(x=c(0,1), y = c(0,1)) +
  theme_net #add direction later
@
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/email_ex-1.pdf}
\caption{\label{fig.cap:email_ex} Email network within a company over a two week period.}
\end{figure}
\afterpage{\clearpage}

@@

  We constructed \texttt{geom\_net} using the drawing capabilities of the \texttt{geom\_point}, \texttt{geom\_line}, \emph{and \texttt{geom\_text}} functions that were already in \texttt{ggplot2}.  Using these capabilities, we were able to seamlessly incorporate our network geometry into the \texttt{ggplot2} structure.   We took our edges data frame and used that as our input to the \texttt{GeomSegment} draw function.  Then, on top of that, we plot the vertices by passing our vertex data frame to the \texttt{GeomPoint} draw function. Finally, we use the \texttt{GeomText} draw function to add the vertex labels when the \texttt{vlabel} argument is TRUE. We added an if/else statement in the \texttt{draw} function which created a labeling layer which is \texttt{NULL} if \texttt{vlabel = FALSE}, which is the default. We then created a new data frame for the label layer. This allows the vertex properties to be propogated through to the labels, including size and color.

>>>>

\subsection{Email Network}
  \emph{Note: would be best with coloring and direction}
  This email network comes from the 2014 VAST Challenge \citep{emailnet}. It is a directed network of emails between company employees with 55 vertices and 9,063 edges. Each vertex is an employee of the company, and each edge is an email sent from one employee to one or more other employees. The arrow of the directed edge points to the recipient(s) of the email. The network contains two business weeks of emails across the entire company. }
<<email_ex, fig.cap = 'All emails within a company over a two week period.'>>=
library(plyr)
email.net <- ddply(email.net, .(eID), transform, nrecipients=length(to))
employee <- read.csv("data/EmployeeRecords.csv")
employee <- merge(email.names, employee, by.x="label", by.y="EmailAddress", all.x=T)
ggplot(data = subset(email.net, nrecipients < 54), aes(from_id = From, to_id = to)) +
  geom_net(vertices = employee,ealpha = 0.1, vsize = I(3), directed = TRUE, aes(vcolour=as.numeric(CurrentEmploymentType))) +
  expand_limits(x=c(0,1), y = c(0,1)) +
  theme_net #add direction later
@
\begin{figure}
\centering
\includegraphics[width=\textwidth]{figure/email_ex-1.pdf}
\caption{\label{fig.cap:email_ex} Email network within a company over a two week period.}
\end{figure}
\afterpage{\clearpage}

@@

  We constructed \texttt{geom\_net} using the drawing capabilities of the \texttt{geom\_point}, \texttt{geom\_line}, \emph{and \texttt{geom\_text}} functions that were already in \texttt{ggplot2}.  Using these capabilities, we were able to seamlessly incorporate our network geometry into the \texttt{ggplot2} structure.   We took our edges data frame and used that as our input to the \texttt{GeomSegment} draw function.  Then, on top of that, we plot the vertices by passing our vertex data frame to the \texttt{GeomPoint} draw function. Finally, we use the \texttt{GeomText} draw function to add the vertex labels when the \texttt{vlabel} argument is TRUE. We added an if/else statement in the \texttt{draw} function which created a labeling layer which is \texttt{NULL} if \texttt{vlabel = FALSE}, which is the default. We then created a new data frame for the label layer. This allows the vertex properties to be propogated through to the labels, including size and color.}
